\begin{thebibliography}{99}

\bibitem{who2021}
World Health Organization, ``Ambient (outdoor) air pollution,'' WHO Fact Sheet, 2021.

\bibitem{iqair2023}
IQAir, ``World Air Quality Report 2023,'' IQAir, 2023.

\bibitem{box1976}
G.~E.~P.~Box and G.~M.~Jenkins, \textit{Time Series Analysis: Forecasting and Control}. Holden-Day, 1976.

\bibitem{hochreiter1997}
S.~Hochreiter and J.~Schmidhuber, ``Long short-term memory,'' \textit{Neural Computation}, vol.~9, no.~8, pp.~1735--1780, 1997.

\bibitem{cho2014}
K.~Cho, B.~van Merrienboer, C.~Gulcehre, D.~Bahdanau, F.~Bougares, H.~Schwenk, and Y.~Bengio, ``Learning phrase representations using RNN encoder-decoder for statistical machine translation,'' in \textit{Proc. EMNLP}, 2014.

\bibitem{vaswani2017}
A.~Vaswani, N.~Shazeer, N.~Parmar, J.~Uszkoreit, L.~Jones, A.~N.~Gomez, L.~Kaiser, and I.~Polosukhin, ``Attention is all you need,'' in \textit{Advances in Neural Information Processing Systems}, 2017.

\bibitem{zhou2021}
H.~Zhou, S.~Zhang, J.~Peng, S.~Zhang, J.~Li, H.~Xiong, and W.~Zhang, ``Informer: Beyond efficient transformer for long sequence time-series forecasting,'' in \textit{Proc. AAAI}, 2021.

\bibitem{wu2021}
H.~Wu, J.~Xu, J.~Wang, and M.~Long, ``Autoformer: Decomposition transformers with auto-correlation for long-term series forecasting,'' in \textit{Advances in Neural Information Processing Systems}, 2021.

\bibitem{yan2018}
S.~Yan, Y.~Xiong, and D.~Lin, ``Spatial temporal graph convolutional networks for skeleton-based action recognition,'' in \textit{Proc. AAAI}, 2018.

\bibitem{chen2016}
T.~Chen and C.~Guestrin, ``XGBoost: A scalable tree boosting system,'' in \textit{Proc. KDD}, 2016.

\bibitem{ke2017}
G.~Ke, Q.~Meng, T.~Finley, T.~Wang, W.~Chen, W.~Ma, Q.~Ye, and T.-Y.~Liu, ``LightGBM: A highly efficient gradient boosting decision tree,'' in \textit{Advances in Neural Information Processing Systems}, 2017.

\bibitem{akiba2019}
T.~Akiba, S.~Sano, T.~Yanase, T.~Ohta, and M.~Koyama, ``Optuna: A next-generation hyperparameter optimization framework,'' in \textit{Proc. KDD}, 2019.

\bibitem{paszke2019}
A.~Paszke, S.~Gross, F.~Massa, A.~Lerer, J.~Bradbury, G.~Chanan, T.~Killeen, Z.~Lin, N.~Gimelshein, L.~Antiga, et al., ``PyTorch: An imperative style, high-performance deep learning library,'' in \textit{Advances in Neural Information Processing Systems}, 2019.

\bibitem{pytorch2019}
PyTorch, ``PyTorch Documentation,'' 2019. [Online]. Available: \url{https://pytorch.org}

\bibitem{onnxruntime2019}
ONNX Runtime, ``ONNX Runtime: cross-platform, high performance ML inferencing and training accelerator,'' Microsoft, 2019. [Online]. Available: \url{https://onnxruntime.ai}

\bibitem{zhang2019}
Y.~Zhang, H.~Li, and X.~Wang, ``Air quality forecasting with hybrid deep learning models,'' \textit{Atmospheric Environment}, vol.~209, pp.~115--123, 2019.

\bibitem{kumar2020}
A.~Kumar and P.~Goyal, ``Forecasting of air quality index in Delhi using neural network based on principal component analysis,'' \textit{Pure and Applied Geophysics}, vol.~177, pp.~2891--2908, 2020.

\bibitem{li2020}
S.~Li, X.~Jin, Y.~Xuan, X.~Zhou, W.~Chen, Y.-X.~Wang, and X.~Yan, ``Enhancing the locality and breaking the memory bottleneck of transformer on time series forecasting,'' in \textit{Advances in Neural Information Processing Systems}, 2019.

\bibitem{kingma2015}
D.~P.~Kingma and J.~Ba, ``Adam: A method for stochastic optimization,'' in \textit{Proc. ICLR}, 2015.

\bibitem{pedregosa2011}
F.~Pedregosa, G.~Varoquaux, A.~Gramfort, V.~Michel, B.~Thirion, O.~Grisel, M.~Blondel, P.~Prettenhofer, R.~Weiss, V.~Dubourg, et al., ``Scikit-learn: Machine learning in Python,'' \textit{Journal of Machine Learning Research}, vol.~12, pp.~2825--2830, 2011.

\bibitem{wes2010}
M.~Wes, ``pandas: a foundational Python library for data analysis and statistics,'' \textit{Python for High Performance and Scientific Computing}, 2011.

\bibitem{seabold2010}
S.~Seabold and J.~Perktold, ``Statsmodels: Econometric and statistical modeling with Python,'' in \textit{Proc. SciPy}, 2010.

\bibitem{cpcd2023}
Central Pollution Control Board, ``National Air Quality Monitoring Programme (NAMP),'' Government of India, 2023.

\bibitem{lane2016}
N.~D.~Lane, S.~Bhattacharya, P.~Georgiev, C.~Forlivesi, L.~Jiao, L.~Qendro, and F.~Kawsar, ``DeepX: A software accelerator for low-power deep learning inference on mobile devices,'' in \textit{Proc. IPSN}, 2016.

\bibitem{wang2020}
Q.~Wang, F.~Li, Y.~Tang, and B.~Yin, ``Edge AI: On-demand accelerating deep neural network inference via edge computing,'' \textit{IEEE Transactions on Wireless Communications}, vol.~19, no.~1, pp.~447--457, 2020.

\bibitem{wen2022}
Q.~Wen, T.~Zhou, C.~Zhang, W.~Chen, Z.~Ma, J.~Yan, and L.~Sun, ``Transformers in time series: A survey,'' in \textit{Proc. IJCAI}, 2022.

\bibitem{benidis2022}
K.~Benidis, S.~S.~Rangapuram, V.~Flunkert, Y.~Wang, D.~Maddix, C.~Turkmen, J.~Gasthaus, M.~Bohlke-Schneider, D.~Salinas, L.~Stella, et al., ``Deep learning for time series forecasting: Tutorial and literature survey,'' \textit{ACM Computing Surveys}, vol.~55, no.~6, 2022.

\bibitem{sharma2020}
E.~Sharma, R.~Chandra, and A.~Mishra, ``Prediction of PM$_{2.5}$ concentration using deep learning models,'' \textit{Environmental Monitoring and Assessment}, vol.~192, 2020.

\bibitem{liu2022}
M.~Liu, A.~Zeng, M.~Chen, Z.~Xu, Q.~Lai, L.~Ma, and Q.~Xu, ``SCINet: Time series modeling and forecasting with sample convolution and interaction,'' in \textit{Advances in Neural Information Processing Systems}, 2022.

\bibitem{openmeteo}
Open-Meteo, ``Free Weather API: Open-Meteo Air Quality API Documentation,'' 2024. [Online]. Available: \url{https://open-meteo.com/en/docs/air-quality-api}

\end{thebibliography}
