% VCE Final Year Project Report
\documentclass[12pt,a4paper]{report}

% Packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{mathptmx}
\usepackage{graphicx}
\usepackage{geometry}
\usepackage{setspace}
\usepackage{titlesec}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{tabularx}
\usepackage{multirow}
\usepackage{enumitem}
\usepackage{listings}
\usepackage{xcolor}
\usepackage{fancyhdr}
\usepackage{hyperref}
\usepackage{url}
\usepackage{float}
\usepackage{tikz}
\usetikzlibrary{positioning,arrows.meta,shapes.geometric,fit,backgrounds,calc,matrix,chains}

% Color definitions for architecture diagrams
\definecolor{lstmblue}{RGB}{70,130,180}
\definecolor{gruorange}{RGB}{220,140,60}
\definecolor{attngreen}{RGB}{60,180,75}
\definecolor{ffpurple}{RGB}{150,80,180}
\definecolor{cnnred}{RGB}{200,70,70}

% Image path
\graphicspath{{./images/}}

% Geometry: 1 inch margins
\geometry{a4paper, left=1in, right=1in, top=1in, bottom=1in}

% Line spacing: 1.5
\onehalfspacing

% Title page and front matter: single spacing
\newcommand{\frontspacing}{\singlespacing}

% Page numbering: bottom center
\pagestyle{fancy}
\fancyhf{}
\fancyfoot[C]{\thepage}
\renewcommand{\headrulewidth}{0pt}
\fancypagestyle{plain}{\fancyhf{}\fancyfoot[C]{\thepage}\renewcommand{\headrulewidth}{0pt}}

% Heading formatting: 14-16pt
\titleformat{\chapter}[display]
  {\normalfont\fontsize{16}{19}\bfseries\centering}
  {\MakeUppercase{\chaptertitlename}\ \thechapter}{12pt}{\fontsize{16}{19}\bfseries\MakeUppercase}
\titleformat{\section}
  {\normalfont\fontsize{14}{17}\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}
  {\normalfont\fontsize{13}{16}\bfseries}{\thesubsection}{1em}{}
\titlespacing*{\chapter}{0pt}{-20pt}{20pt}

% Code listing style
\definecolor{codebg}{rgb}{0.95,0.95,0.95}
\lstset{
    backgroundcolor=\color{codebg},
    basicstyle=\ttfamily\small,
    breaklines=true,
    frame=single,
    numbers=left,
    numbersep=5pt,
    tabsize=4
}

% Hyperref setup
\hypersetup{
    colorlinks=true, linkcolor=black, citecolor=black, urlcolor=blue,
    pdftitle={Comprehensive Air Quality Forecasting and Edge AI Deployment}
}

% Title page info
\newcommand{\projecttitle}{COMPREHENSIVE AIR QUALITY FORECASTING AND EDGE AI DEPLOYMENT}
\newcommand{\studentA}{Kotichintala Srihari Sakshith}
\newcommand{\htnoA}{1602-22-748-046}
\newcommand{\studentB}{Kamal Bura}
\newcommand{\htnoB}{1602-22-748-011}
\newcommand{\studentC}{Mohammad Abdul Muttalib}
\newcommand{\htnoC}{1602-22-748-307}
\newcommand{\guide}{Sriperambuduri Vinay Kumar}
\newcommand{\designation}{Assistant Professor}
\newcommand{\hod}{Dr. T. Adilakshmi}
\newcommand{\dept}{Department of Computer Science \& Engineering}
\newcommand{\college}{Vasavi College of Engineering (Autonomous)}
\newcommand{\affiliation}{(Affiliated to Osmania University)}
\newcommand{\location}{Ibrahimbagh, Hyderabad-31}
\newcommand{\theyear}{2025}

\begin{document}

% Title Page
\begin{titlepage}
\singlespacing
    \centering
    \vspace*{0.2in}
    {\Large \textbf{\college} \par}
    \vspace{0.1in}
    {\large \textbf{\affiliation} \par}
    \vspace{0.1in}
    {\large \location \par}
    \vspace{0.4in}
    {\Large \textbf{\dept} \par}
    \vspace{0.4in}
    {\LARGE \textbf{\projecttitle} \par}
    \vspace{0.3in}
    {\large A Project Report \par}
    {\large Submitted in partial fulfillment of the requirements \par}
    {\large for the award of the degree of \par}
    \vspace{0.2in}
    {\LARGE \textbf{BACHELOR OF ENGINEERING} \par}
    {\large \textbf{IN} \par}
    {\LARGE \textbf{COMPUTER SCIENCE \& ENGINEERING} \par}
    \vspace{0.4in}
    {\large \textbf{Submitted by} \par}
    \vspace{0.2in}
    {\large \textbf{\studentA} \par}
    {\large (\htnoA) \par}
    \vspace{0.1in}
    {\large \textbf{\studentB} \par}
    {\large (\htnoB) \par}
    \vspace{0.1in}
    {\large \textbf{\studentC} \par}
    {\large (\htnoC) \par}
    \vspace{0.4in}
    {\large \textbf{Under the Guidance of} \par}
    \vspace{0.1in}
    {\large \textbf{\guide} \par}
    {\large \designation \par}
    \vfill
    {\large \textbf{\dept} \par}
    {\large \textbf{\college} \par}
    {\large \textbf{\theyear} \par}
\end{titlepage}

% Certificate
\newpage\thispagestyle{empty}\singlespacing
\begin{center}
    {\Large \textbf{\college} \par}
    \vspace{0.08in}
    {\large \affiliation \par}
    \vspace{0.08in}
    {\large Hyderabad - 500 031 \par}
    \vspace{0.2in}
    {\Large \textbf{\dept} \par}
    \vspace{0.3in}
    {\Large \textbf{BONAFIDE CERTIFICATE} \par}
\end{center}
\vspace{0.3in}
\hspace{0.5in} This is to certify that the project entitled \textbf{``\projecttitle''} being submitted by \textbf{\studentA} (\htnoA), \textbf{\studentB} (\htnoB) and \textbf{\studentC} (\htnoC), in partial fulfillment of the requirements for the award of the degree of \textbf{Bachelor of Engineering in Computer Science \& Engineering} is a record of bonafide work carried out by them under my guidance and supervision.

\vspace{0.4in}
\begin{flushright}
    \textbf{\guide} \\
    \designation \\
    Department of CSE \\
    \vspace{0.1in}
    \textbf{Internal Guide}
\end{flushright}

\vspace{0.4in}
\begin{flushleft}
    \textbf{\hod} \\
    Professor \& HOD \\
    Department of CSE
\end{flushleft}

\vspace{0.4in}
\begin{center}\textbf{External Examiner}\end{center}

% Declaration
\newpage\thispagestyle{empty}\singlespacing
\begin{center}
    {\Large \textbf{DECLARATION} \par}
\end{center}
\vspace{0.3in}
We, \textbf{\studentA} (\htnoA), \textbf{\studentB} (\htnoB) and \textbf{\studentC} (\htnoC), students of Computer Science \& Engineering, \college, Hyderabad hereby declare that the project report entitled \textbf{``\projecttitle''} submitted to the Department of Computer Science \& Engineering, \college, Hyderabad, is a record of authentic work carried out by us under the supervision of \textbf{\guide}, \designation, Department of CSE.

\vspace{0.3in}
We further declare that this project work has not been submitted previously for the award of any degree or diploma in any other university or institution.

\vspace{0.8in}
\begin{flushright}
    \textbf{\studentA} \\
    (\htnoA) \\
    \vspace{0.3in}
    \textbf{\studentB} \\
    (\htnoB) \\
    \vspace{0.3in}
    \textbf{\studentC} \\
    (\htnoC)
\end{flushright}
\vspace{0.4in}
\begin{flushleft}
    Date: \underline{\hspace{1in}} \\
    Place: Hyderabad
\end{flushleft}

% Acknowledgement
\newpage\thispagestyle{empty}\singlespacing
\begin{center}{\Large \textbf{ACKNOWLEDGEMENT}\par}\end{center}
\vspace{0.3in}
We express our sincere gratitude to all those who have contributed to the successful completion of this project.

\vspace{0.2in}
We are deeply indebted to our project guide, \textbf{\guide}, \designation, Department of Computer Science \& Engineering, for his invaluable guidance, constant encouragement, and constructive suggestions throughout the course of this project.

\vspace{0.2in}
We extend our heartfelt thanks to \textbf{\hod}, Professor and Head of the Department of Computer Science \& Engineering, for providing us with the necessary facilities and resources to carry out this project work.

\vspace{0.2in}
We are grateful to \textbf{Dr. S. V. Ramana}, Principal, \college, for his administrative support and for creating an excellent academic environment.

\vspace{0.2in}
We also thank all the faculty members of the Department of Computer Science \& Engineering for their continuous support and encouragement. We extend our appreciation to our classmates and friends who provided valuable feedback and assistance during the development of this project.

\vspace{0.2in}
Finally, we thank our parents and family members for their unconditional love, support, and motivation throughout our academic journey.

\vspace{0.6in}
\begin{flushright}
    \textbf{\studentA} \\
    \textbf{\studentB} \\
    \textbf{\studentC}
\end{flushright}

% Abstract
\newpage\thispagestyle{empty}\singlespacing
\begin{center}{\Large \textbf{ABSTRACT}\par}\end{center}
\vspace{0.3in}
Air pollution poses significant health risks to urban populations, making accurate air quality forecasting essential for public health management and policy formulation. This project presents a comprehensive framework for air quality forecasting using state-of-the-art machine learning and deep learning models, with deployment on resource-constrained edge devices.

We systematically evaluate 14 forecasting models spanning six architectural categories: statistical baselines (ARIMA, SARIMA, VAR), standard deep learning sequence models (RNN, LSTM, GRU, BiLSTM), hybrid architectures (CNN-LSTM, CNN-GRU, BiLSTM with Attention), transformer-based models (Transformer, Informer, Autoformer), and spatio-temporal networks (ST-GCN). Each model undergoes rigorous hyperparameter optimization using Optuna with 20 trials.

Experimental evaluation on the Hyderabad air quality dataset (8,784 hourly observations sourced from the Open-Meteo API) reveals that the Transformer architecture achieves the best overall performance with an RMSE of 17.34 and $R^2$ of 0.471. The CNN-LSTM hybrid follows closely with RMSE of 17.50, while GRU --- the most efficient deep learning model --- achieves an RMSE of 17.62 with only 38 seconds of training time. Statistical baselines (ARIMA, SARIMA, VAR) demonstrate poor performance with negative $R^2$ values, confirming that simple univariate statistical approaches are inadequate for the complex multivariate dynamics of PM$_{2.5}$ forecasting.

We further demonstrate edge deployment feasibility by implementing all models on a Raspberry Pi 4 for real-time inference. The system achieves end-to-end latency under 200 ms per prediction with annual operating costs under Rs.~250, establishing a practical framework for distributed, low-cost air quality monitoring networks.

\noindent \textbf{Keywords:} Air Quality Forecasting, Edge AI, Raspberry Pi, Deep Learning, Transformers, LSTM, GRU, PM$_{2.5}$ Prediction, Time Series Analysis, Model Benchmarking.

% Table of Contents, List of Figures, List of Tables
\newpage\thispagestyle{plain}\tableofcontents
\newpage\thispagestyle{plain}\listoffigures
\newpage\thispagestyle{plain}\listoftables

% Chapters
\input{ch1_introduction}
\input{ch2_system_analysis}
\input{ch3_system_design}
\input{ch4_implementation}
\input{ch5_results}
\input{ch6_conclusion}
\input{ch7_references}

\end{document}
